---
title: "merge_AHRQ"
output: html_notebook
---

# Load libraries
```{r}
library(caret)
library(data.table)
library(dplyr)
library(sets)
library(scales)
library(tidyr)
library(stringr)
```

# Load patient data:
```{r}
# Import (synthetic) patient dataset
df_patients <- read.csv("../Data/fake_patient_data.csv", header = TRUE)

head(df_patients)
```

# Load ZIP code level AHRQ SDOHD Data
```{r}
# Import AHRQ data
zip_ahrq_df_2011 <- read.csv("../Data/AHRQ/ZIP/sdoh_2011_ZIPCODE_1_0.csv")
zip_ahrq_df_2012 <- read.csv("../Data/AHRQ/ZIP/sdoh_2012_ZIPCODE_1_0.csv") # header=FALSE
zip_ahrq_df_2013 <- read.csv("../Data/AHRQ/ZIP/sdoh_2013_ZIPCODE_1_0.csv") # header=FALSE
zip_ahrq_df_2014 <- read.csv("../Data/AHRQ/ZIP/sdoh_2014_ZIPCODE_1_0.csv") # header=FALSE
zip_ahrq_df_2015 <- read.csv("../Data/AHRQ/ZIP/sdoh_2015_ZIPCODE_1_0.csv") # header=FALSE
zip_ahrq_df_2016 <- read.csv("../Data/AHRQ/ZIP/sdoh_2016_ZIPCODE_1_0.csv") # header=FALSE
zip_ahrq_df_2017 <- read.csv("../Data/AHRQ/ZIP/sdoh_2017_ZIPCODE_1_0.csv") # header=FALSE
zip_ahrq_df_2018 <- read.csv("../Data/AHRQ/ZIP/sdoh_2018_ZIPCODE_1_0.csv") # header=FALSE
zip_ahrq_df_2019 <- read.csv("../Data/AHRQ/ZIP/sdoh_2019_ZIPCODE_1_0.csv") # header=FALSE
zip_ahrq_df_2020 <- read.csv("../Data/AHRQ/ZIP/sdoh_2020_ZIPCODE_1_0.csv") # header=FALSE


# Use fill to ensured consistent number of columns, merge dfs from 2009-2020 (via mean).
# Merge dfs on the 6 items below in list.
zip_ahrq_total <- rbindlist(list(zip_ahrq_df_2011, zip_ahrq_df_2012, zip_ahrq_df_2013,
                             zip_ahrq_df_2014, zip_ahrq_df_2015, zip_ahrq_df_2016, zip_ahrq_df_2017, zip_ahrq_df_2018,
                             zip_ahrq_df_2019, zip_ahrq_df_2020), fill=TRUE)
```


# AHRQ SDOH Feature names: at the ZIP code level. 
- this file has the *union* of all feature names from 2011-2020. 
- For each variable, this file contains: 
    - Variable_name: unique variable identifier
      - Variable_name_no_ZC_suffix: the Variable_name field without the suffix "ZC" for zip code.
    - Data_source: the data source, 
    - Domain: SDOH domain, 
    - SDOH_topic: more specific category than the domain,
    - Variable_Label: short written explanation of the variable.
    
## Note: 
- variables with 'NA' for a field simply describe the geography -- they are to be used for merging with patient data (below), not as SDOH variables.
```{r}

zip_ahrq_feat_names <- read.csv("../Data/AHRQ/ZIP/zip_ahrq_feat_names.csv")
print(zip_ahrq_feat_names)

# Drop columns with 'Identifier' or '6. Geography' for Domain.
zip_ahrq_feat_names <- subset(zip_ahrq_feat_names, !(Domain %in% c("Identifier", "6. Geography")))
print(zip_ahrq_feat_names)


```
# Data Summary: categorical
```{r}

barchart(zip_ahrq_feat_names$Data_source)
barchart(zip_ahrq_feat_names$Domain)
barchart(zip_ahrq_feat_names$SDOH_Topic)

```

# Preprocess AHRQ data:
```{r}

# Pad ZIP codes with leading 0's: so all are exactly 5 digits.
zip_ahrq_total$ZIPCODE <- as.integer(zip_ahrq_total$ZIPCODE)
zip_ahrq_total$ZIPCODE <- sprintf("%05d", zip_ahrq_total$ZIPCODE)
zip_ahrq_total$ZIPCODE <- as.integer(zip_ahrq_total$ZIPCODE)
# TODO: if needed, pad ZIP code in patient data with leading zeroes.

# Pad STATEFIPS codes with leading 0's: so all are exactly 2 digits.
# First, convert to character (instead of <dbl>), then store as string.
zip_ahrq_total$STATEFIPS <- as.integer(zip_ahrq_total$STATEFIPS)
zip_ahrq_total$STATEFIPS <- sprintf("%02d", zip_ahrq_total$STATEFIPS)
zip_ahrq_total$STATEFIPS <- as.integer(zip_ahrq_total$STATEFIPS)
# TODO: if needed, pad State FIPS code in patient data with leading zeroes.

# Perform any imputation if desired
# ex.) can impute AHRQ variables at: geographic levels (within state of GA, or entire USA), across time (2011-2020) or one year at a time.

```


# Merge AHRQ with patient data (example):
- Crosswalk variables: use these to merge with patient dataset.
  - STATEFIPS
  - ZIPCODE
  - YEAR
Example: merge 'df_patients' <- zip_ahrq_total
```{r}

merged_ahrq_patients <- merge(zip_ahrq_total, df_patients, by.x=c("ZIPCODE", "STATEFIPS", "YEAR"), by.y=c("ZIP_CODE", "STATE_FIPS", "baselineyear"))

head(merged_ahrq_patients)

```


# Write merged data to output file:
```{r}

write.csv(merged_ahrq_patients, file="./merged_zip_ahrq_patients_final.csv", row.names=FALSE) 

```
